Past few years have been quite a journey.
I have learned how to do science and I have met amazing people in the process.
First, I want to thank David for none of this would have been possible without him.
He thought me how to think like a scientist and afforded me the space to act like one.
Second, I must express my absolute gratitude for the incredible group of labmates with whom I had the chance to work.
In particular, Maria-Teresa is a wonderful human, I learned from her more than I can reckon and I am forever thankful.
Special thanks should also be extended to Jordane.
He meticulously carried out the experiments and assisted in many aspects of this project.
Corane as well greatly helped with final steps of the experiments.
I am extremely grateful for her efforts.
Furthermore, I sincerely appreciate Anna Montagnini, Ingrid Bureau, J\'{e}r\^{o}me Epsztein, and Julie Koenig for their constructive comments and invaluable input throughout this project and my thesis work.
\par
I also had the pleasure to get to know many colleagues at Inmed who became my friends, were there for me and made my time in Marseille as memorable as it was.
Thanks to Sanaz, for her countless tips that helped me settle here and for her friendship to this date.
To Anass, for numerous coffee breaks during which we discussed literally, everything!
To my delightful officemates, Ludovic who helped me a great deal in early stages of this work, and Stefania who did her best to teach me Italian and proofread this manuscript.
And last but not least, to all my friends, Carla, Romain, Davide, Simona, Lexi, Alla, and Shrisha whose company I cherished everyday.
Their presence made a difference.