\chapter{Introduction}
\label{ch:intro:intro}

For humans and other animals, in any context, adaptive behavior is defined as executing an action that would maximize immediate or future rewards, while minimizing energy expenditure.
For example, consider any solitary hunter that would wait for the right time to attack:
    when the prey is most distracted or vulnerable.
However, in some situations, the appearance of a sensory cue will not only indicate which action should be performed, but also how long, after the appearance of the cue, this action must be initiated~\cite{Balsam2009Map, Nobre2018NatRevNeurosci}.
For instance, athletes performing sprint races learn by experience that the \textit{go} command, signaling them to start running, will be given in two seconds after the \textit{set} command.
False starts, i.e., beginning to run just before the \textit{go} command, demonstrate that athletes estimated the~2~second interval between the \textit{set} and \textit{go} commands.
More generally, the ability of animals to exploit temporal regularities in nature is crucial for survival:
    the appearance of a sensory cue at a given time can predict food availability, predator attack, or mating opportunity~\cite{Kacelnik2002,Gallistel1990book}.
\par
In this chapter, first I introduce the taxonomy used in the timing literature to establish a reference for this manuscript.
Then I discuss two possible mechanisms that could give rise to the perception of elapsed time.
Next, I will review some evidence of how either mechanism is thought to be implemented in the brain, focusing on the role of a certain brain structure of interest.
Finally, I will lay out the question and the hypothesis underlying this work and the structure of the following chapters.

\input{ch-intro/intro-topic0}
\input{ch-intro/intro-topic1}
\input{ch-intro/intro-topic2}
\input{ch-intro/intro-topic3}
\input{ch-intro/intro-topic4}