\section{Technical Tools} \label{ch:methods:tech}

\subsection{Striatal Lesion} \label{ch:method:lesion}
\input{ch-methods/figures/LesionSizeLocation.tex}
Anesthesia was induced with an intraperitoneal (IP) injection of a mixture of 100~mg/kg ketamine and 10~mg/kg xylazine and was maintained during the surgery with inhalant isoflurane gas (less than 3\%).
After shaving and cleaning the scalp, the animal was placed in the stereotaxic frame (Kopf instruments) and a local anesthetic (lidocaine) was injected under the scalp.
Then, an incision along the midline of the skull was made, followed by cleaning the exposed skull and drilling the craniotomies above the targeted areas.
To perform fiber-sparing lesion of the \gls{ds}, ibotenic acid (1\% in 0.1~M~NaOH, Fisher Scientific) was infused (Pump 11 Elite Nanomite, Harvard Apparatus, using a 10~$\mu$L WPI Nanofil syringe) in 6 specular sites bilaterally, at a rate of 90~nL/min.
The needle remained in place for 10~min following the injection to allow for the diffusion of the excitotoxic drug.
Then, the needle was retracted slowly to avoid backflow of the drug.
Once all the injections were performed, craniotomies were filled with bone wax, the skull was disinfected, and the skin was sutured.
Animals were allowed to recover for two weeks before resuming behavioral training.
After surgery, animals were housed alone for 3 days, to avoid getting hurt by the cagemates, and were force-fed if needed.
Injection coordinates (in~mm, with reference to Bregma, according to Paxinos) are shown in \Autoref{fig:method:LesionSizeLocation}{A} (each injection at $-5.6$~mm dorsoventral).
The infused volume in each site was 200~nL for DLS and DMS lesions, and 400~nL for \gls{ds} lesions.

\subsection{Immunohistochemistry}
At the end of the experiments, animals were euthanized with an overdose IP injection of $\sim$2~mL pentobarbital.
Then, they were perfused with 4\% paraformaldehyde and their brains were harvested for histological analysis of the lesion size and location.
Brains were coronally sliced on a vibratome at 60~$\mu$m thickness.
For each animal, six sections spanning the \gls{ds} along the rostrocaudal axis were selected (usually the following slice numbers: 5, 15, 25, 35, 45, and 55 for consistency) and submerged in 0.1~M~PBS.
Then, PBS was replaced with citrate buffer (10~mM) for 10~min at room temperature.
Next, slices were submerged with a blocking solution, consisting of PBS with 0.3\% triton and 15\% normal goat serum (NGS) for 120~min at room temperature.
Then, the solution was replaced with another consisting of 2~$\mu$L anti-NeuN antibody (Merks Millipore, MAB377) and 0.5~$\mu$L of anti-GFAP antibody (Agilent, Z033429-2) diluted in 200~$\mu$L of the blocking solution, kept overnight at 4~$^{\circ}$C.
Sections were then rinsed twice in PBS, 10~min each, at room temperature, before being submerged again in 1~$\mu$L of donkey anti-mouse antibody (Al555, red), 1~$\mu$L of donkey anti-rabbit antibody (AL488, green) diluted in 400~$\mu$L of PBS for 120~min at room temperature.
Finally, they were washed twice in PBS, for 10~min each time, and mounted for microscopy.

\subsection{Lesion Quantification}
Whole slices were imaged using an Apotome microscope (Zeiss, 28126), and stitched together in the processing software (Zeiss Zen).
Then for each slice, the ventricule, the striatum, and the lesioned area were manually outlined (\Autoref{fig:method:LesionSizeLocation}{B}) bilaterally in the image processing software (ImageJ, Fiji).
The size and the centroid coordinates were automatically computed for all of the above-mentioned areas.
Next, the anteroposterior location of each slice was also approximated according to the rat brain atlas (Paxinos).
\par
The lesion size reported in this paper is the ratio of the lesion volume over the volume of the striatum.
Both regions of interest (lesion and striatum) were approximated as a truncated cone between any two consecutive sections, and the volume was accordingly calculated and summed up.
\Autoref{fig:method:LesionSizeLocation}{C-D} show the lesion coordinates of all the animals trained in the normal treadmill task.
Four rat brains were imaged improperly, they are automatically removed from any analysis that required the lesion size.


\subsection{Statistics}
All statistical comparisons were performed using a permutation test previously described in~\cite{Fujisawa2008NN}.
This non-parametric method alleviates many concerns in statistical hypothesis tests, such as distribution assumptions (e.g., normality assumption under analysis of variance), error inflation due to multiple comparisons, and sensitivity to unbalanced group size.
\par
To simplify the description, let's assume, we have ${\mathbf{X}=[X_1, X_2,...,X_n]}$, where $X_i$ is the set of $ET$s of session~$i$.
Similarly, we have $\mathbf{Y}$ that contains \glspl{et} of all the sessions from another condition.
Here, the null hypothesis states that the assignment of each data point in $X_i$ and $Y_i$ to either $\mathbf{X}$ or $\mathbf{Y}$ is random, hence there is no difference between $\mathbf{X}$ and $\mathbf{Y}$.
\par
In short, the test statistic was defined as the difference between smoothed (using Gaussian kernel with $\sigma =0.05$) $\mathbf{X}$ and $\mathbf{Y}$ for each session~$i$: $D_0(i)$.
At this point, we generated one set of surrogate data by assigning each \gls{et} of session $i$ to either $X_i$ or $Y_i$, randomly.
For each set of surrogate data, the test statistic was calculated, i.e.,~$D_m(i)$.
This process was repeated 10,000 times for all the statistical comparisons in this study, obtaining: $D_1(i),\ldots,D_{10000}(i)$.
\par
At this step, two-tailed pointwise p-values could be directly calculated for each $i$, from the $D_m(i)$ quantiles~\cite[see][]{Fujisawa2008NN}.
Moreover, to compensate for the issue of multiple comparisons, we defined global bands of significant differences along the session index dimension.
From 10,000 sets of surrogate data, band of the largest $\alpha$-percentile was constructed, such that less than 5\% of $D_m(i)$s broke the band at any given session $i$.
This band (denoted as the \textit{global band}) represents the threshold for significance, and any break-point by $D_0(i)$ at any $i$ is a point of significant difference between $\mathbf{X}$ and $\mathbf{Y}$.
\par
In cases of comparing only two sets of data points, the same algorithm was employed, having only one value for index $i$.
If none of the $D_m(i)$s exceeded $D_0(i)$, the value $p<0.0001$ was reported (i.e., less than one chance in 10,000).
